Emulsions Stabilized by Magnetic Ellipsoidal Particles: A Lattice Boltzmann Study

### 2025-04-08 chapters/LBM_technique.tex
Changes 1. Added new lit review 2. Refined introduction and LBM methods 3. Fixed all references so that they have journals

\section{Methods}
\label{section:methods}

@@

For an incompressible fluid—meaning constant density and negligible bulk viscosity—the condition $\nabla \cdot \vec{u} = 0$ holds. Under this assumption, the Navier-Stokes equation simplifies to,

@@

@@

\begin{equation}
    \frac{\partial \bar{\vec{u}}}{\partial \bar{t}} + (\bar{\vec{u}} \cdot\nabla)\bar{\vec{u}} = -\nabla \bar{p} + \frac{1}{Re} \Delta \bar{\vec{u}} + \bar{\vec{f}}    
\end{equation}

@@

\section{Phase separation}
Another physical phenomenon relevant to bijel formation is non-ideal mixing. The free energy of mixing for a binary fluid is given by the thermodynamic relation

\begin{equation}
    \Delta G_m = \Delta H_m - T \Delta S_m
\end{equation}

In an ideal solution, the enthalpy of mixing is zero, i.e., $\Delta H_m = 0$, and the entropy of mixing $\Delta S_m$ can be derived using statistical mechanical approximations of the 
configurational states of the binary mixture. For a system with volume fractions $\phi$ and $1 - \phi$, the entropy of mixing becomes

\begin{equation}
    \Delta S_m = -N_a k_B \left[(1 - \phi)\ln(1 - \phi) + \phi \ln(\phi)\right]
\end{equation}

In contrast, a regular solution includes non-zero enthalpic interactions. The total free energy of mixing is split into ideal and excess components,

\begin{equation}
    \Delta G_m = \Delta G_i + \Delta G_e
\end{equation}

The ideal contribution arises from the entropy of mixing: $\Delta G_i = -T \Delta S_m$. For regular solutions, the excess entropy is assumed to be zero, and the excess enthalpy is given by,

\begin{equation}
    \Delta H_m = \chi (1 - \phi)\phi
\end{equation}

where $\chi$ is the interaction parameter quantifying the energetic favorability of mixing. This term originates from mean-field approximations assuming equal-sized molecular species. 
Combining both contributions, the free energy of mixing becomes:

\begin{equation}
    \Delta G_m = \chi (1 - \phi)\phi + N_a k_B T \left[(1 - \phi)\ln(1 - \phi) + \phi \ln(\phi)\right]
\end{equation}

This bulk free energy characterizes the thermodynamic driving forces for phase separation. Alternative forms, such as the Landau-Ginzburg or Flory-Huggins models, may also be used depending on system 
specifics. In real systems, the interface between two partially miscible fluids is not sharp; rather, the composition transitions smoothly across a finite interface width. To account for this, a 
gradient energy penalty is added to the free energy functional, penalizing sharp spatial variations in composition. The total free energy functional is written as

\begin{equation}
    F[\phi] = \int \left( f(\phi) + \frac{\kappa}{2}|\nabla \phi|^2 \right) \, dV
\end{equation}

where $\kappa$ is the gradient energy coefficient controlling the interface width. Larger \(\kappa\) values correspond to wider interfaces.
The chemical potential is then defined as the variational derivative of the free energy functional,

\begin{equation}
    \mu = \frac{\delta F}{\delta \phi} = \frac{df}{d\phi} - \kappa \nabla^2 \phi
\end{equation}

This chemical potential enters the mass flux expression via a Fickian transport law,

\begin{equation}
    \vec{J} = -M \nabla \mu
\end{equation}

where $M$ is the mobility and could potentially be a function of $\phi$. Assuming no advection and applying mass conservation for the order parameter, we obtain,

\begin{equation}
    \frac{\partial \phi}{\partial t} = -\nabla \cdot \vec{J}
\end{equation}

Substituting in the expression for \(\vec{J}\) leads to the Cahn-Hilliard equation,

@@

This equation describes the diffusive evolution of the conserved field $\phi$ during phase separation. It captures the system's tendency to minimize its total free energy 
by redistributing the order parameter according to gradients in chemical potential. The Cahn-Hilliard framework naturally incorporates interfacial tension and allows for smooth, 
continuous transitions between phases, making it highly suitable for modeling non-ideal mixing in systems such as bijels.

@@

Ultimately, the choice between point and explicit particle models reflects a trade-off between computational efficiency and physical realism. For systems in which interfacial structure, 
jamming, or near-field hydrodynamics play a central role as in bijel formation, explicit particle modeling is generally preferred.

@@

To address these limitations, the demagnetization tensor method offers a more refined model by considering the particle as a uniformly magnetized ellipsoid and calculating the internal magnetic 
field using an analytical or numerically tabulated demagnetization tensor. \cite{takahashi_ellipsoids_2017} 
This method accounts for the effect of particle shape on the distribution of internal magnetization and enables the modeling of induced magnetic dipoles that respond to local fields. 
The demagnetization field \(\vec{H}_d\) inside a particle is given by

\begin{equation}
    \vec{H}_d = -\mathbf{N} \cdot \vec{M}
\end{equation}

where \(\mathbf{N}\) is the demagnetization tensor and \(\vec{M}\) is the magnetization vector. \cite{takahashi_ellipsoids_2017} 
This approach is particularly useful for modeling spheroidal or ellipsoidal particles and is widely used in magnetorheological systems, chain-forming ferrofluids, and dense suspensions where 
particle-particle interactions are strongly shape-dependent.

At a higher level of fidelity, micromagnetic continuum models resolve the internal magnetization field \(\vec{M}(\vec{r})\) throughout each particle by minimizing the total magnetic energy, which 
includes exchange, anisotropy, Zeeman, and magnetostatic contributions. \cite{vivas_toward_2020} 
The governing equations are typically derived from the Landau-Lifshitz-Gilbert (LLG) equation, which evolves \(\vec{M}(\vec{r}, t)\) over time under the influence of effective magnetic fields. 
These models capture spatially varying domain structures, vortex states, and other complex magnetic configurations, making them essential for accurately simulating particles with nonuniform or dynamic 
magnetization. However, micromagnetic simulations are computationally expensive and require fine discretization within each particle, often using finite element or finite difference methods.

@@

The Volume of Fluid (VOF) method is a sharp-interface technique used to track the evolution of fluid-fluid interfaces within a fixed computational grid.
\cite{gopala_volume_2008, fleckenstein_volume-fluid-based_2015, deising_direct_2018}
In VOF, a scalar field (the volume fraction) is defined for each grid cell to indicate the presence of each fluid component, with interface reconstruction algorithms used to localize the boundary 
between phases. 
The method conserves mass by construction and handles large interface deformations well, making it suitable for high-Reynolds-number flows and surface tension-driven dynamics. However, 
VOF methods can struggle with accurately capturing interfacial curvature and do not easily accommodate diffuse interfaces or complex topological changes such as pinch-off or coalescence,
which limits their applicability for simulating phase separation in bijels.

Phase field methods take a diffuse-interface approach, representing fluid composition or phase identity using a continuous order parameter field and are often governed by the Cahn-Hilliard or 
Allen-Cahn equations. \cite{mendoza_evolution_2006, carmack_tuning_2018, chan_channel_2012}
These methods naturally capture smooth transitions between phases, allow for accurate calculation of interfacial forces, and can easily handle topological changes in the interface. 
Phase field methods are thermodynamically consistent, as they are derived from free energy functionals, and are particularly well-suited for modeling systems undergoing spinodal decomposition 
or wetting transitions. Their main limitations lie in the need for fine resolution near interfaces to maintain numerical accuracy and stability, and in the computational cost of solving 
higher-order partial differential equations, especially when coupled with full hydrodynamics.

Brownian dynamics is a particle-based method used to simulate the stochastic motion of colloidal particles in a fluid. 
\cite{huber_brownian_2019, yip_brownian_2005, elsawy_utility_2025}
It models the over-damped limit of Langevin dynamics by integrating particle 
trajectories under the influence of deterministic forces (such as interparticle interactions or external fields) and random thermal forces representing collisions with solvent molecules. Hydrodynamic 
interactions may be included via approximate mobility tensors or neglected altogether, depending on the level of detail required. Brownian dynamics is ideal for studying self-assembly, aggregation, 
and diffusion of particles in dilute systems, but lacks the ability to capture bulk fluid motion or interface-mediated interactions unless coupled with continuum solvers. As such, it is not well-suited 
for simulating the coupled fluid-particle dynamics characteristic of bijel formation on its own.

@@

In summary, the Volume of Fluid method offers robust interface tracking and mass conservation but lacks the ability to naturally capture interfacial curvature and topological changes. 
Phase field methods provide thermodynamically consistent, diffuse-interface models ideal for phase separation, though they require fine resolution and are computationally intensive. 
Brownian dynamics excels at modeling particle-level stochastic motion but cannot resolve bulk fluid dynamics or interface evolution on its own. In contrast, the Lattice Boltzmann Method 
combines the strengths of mesoscopic fluid modeling with efficient handling of complex interfaces and straightforward coupling to particle dynamics. Its balance between computational 
efficiency, scalability, and physical fidelity across multiple scales makes LBM the most suitable method for simulating the rich interplay of hydrodynamics, phase separation, and particle 
interactions in stimuli-responsive emulsions such as bijels.

@@

The IT model employs level-set or front-tracking techniques to explicitly track the location of fluid interfaces by evolving a separate indicator 
function alongside the LB solver. \cite{haghani_hassan_abadi_conservative_2018, liang_lattice_2023}
This method achieves high accuracy in resolving interface position and curvature, making it ideal for problems 
involving surface tension-driven flows or sharp interfacial features. However, it is computationally intensive and less well-suited for handling 
topological changes such as coalescence or pinch-off, which are common in phase-separating systems like bijels.

@@

>>>>

@@

For an incompressible fluid, meaning that the constant density and negligible bulk viscosity, the condition $\nabla \cdot \vec{u} = 0$ holds. Under this 
assumption, the Navier-Stokes equation simplifies to,

@@

As there are no external forces like gravity acting on the fluid and there are is no hydrodynamic pressure being applied to the fluid, $-\nabla p = \vec{f} = 0$ resulting in the
following form of the incompressible Navier Stokes equation.

\begin{equation}
    \rho \left(\frac{\partial\vec{u}}{\partial t} + (\vec{u}\cdot\nabla)\vec{u} \right) = \eta \Delta \vec{u}
\end{equation}

@@

\begin{equation}
    \frac{\partial \bar{\vec{u}}}{\partial \bar{t}} + (\bar{\vec{u}} \cdot\nabla)\bar{\vec{u}} = \frac{1}{Re} \Delta \bar{\vec{u}}
\end{equation}

@@

\section{Phase separation}

Bijels form through spinodal decomposition of two partially miscible fluids, a process that can be effectively modeled using the Cahn-Hilliard equation, which describes the diffusive dynamics 
of a conserved scalar order parameter $\phi$, typically representing the local concentration difference between the two fluid components,

@@

Here, $M$ is the mobility coefficient and $\mu$ is the chemical potential, which is derived from the variation of the free energy functional $F[\phi]$. This formulation allows for the modeling of 
interfacial tension and the spontaneous formation of bicontinuous domains characteristic of spinodal decomposition.
To capture not only the diffusive separation of phases but also the resulting fluid motion, the Cahn-Hilliard equation is coupled with the Navie-Stokes equations, leading to the 
Cahn-Hilliard-Navier-Stokes (CHNS) framework. This coupling is essential in systems like bijels, where hydrodynamic interactions play a significant role in domain coarsening and morphology arrest. 
The momentum conservation equation in this framework is written as:

\begin{equation}
    \rho \left(\frac{\partial\vec{u}}{\partial t} + (\vec{u}\cdot\nabla)\vec{u} \right) = \eta \Delta \vec{u} + \nabla \cdot P_{\text{therm}}
\end{equation}

Here, $\vec{u}$ is the velocity field, $\rho$ is the fluid density, and $\eta$ is the dynamic viscosity. The term $P_{\text{therm}}$ represents the thermodynamic pressure tensor, which incorporates 
contributions from gradients in the order parameter and is directly linked to the chemical potential \(\mu\). This tensor encodes the capillary and interfacial forces arising from phase separation 
and effectively couples the compositional dynamics to the fluid flow.
The CHNS model enables simulation of complex phase-separating fluid systems, capturing both the interfacial evolution and the momentum transfer within the fluid. In the context of bijels, this 
model can be used to predict the formation and stabilization of the bicontinuous structure, especially when interfacial jamming by colloidal particles is introduced in more advanced extensions.

@@

For systems in which interfacial structure, jamming, or near-field hydrodynamics play a central role as in bijel formation, explicit particle modeling 
is generally preferred as we are able to resolve the interfacial structure and near field hydrodynamics without the complications of IBM.

@@

To address these limitations, the demagnetization tensor method offers a more refined model by considering the particle as a uniformly magnetized ellipsoid and calculating the internal magnetic 
field using an analytical or numerically tabulated demagnetization tensor. \cite{takahashi_ellipsoids_2017} 
This method accounts for the effect of particle shape on the distribution of internal magnetization and enables the modeling of induced magnetic 
dipoles that respond to local fields. The total magnetic field of an ellipsoids can be written,

\begin{equation}
    \mathbf{H}(r) = \mathbf{H}_0 + \mathbf{N}(r)\mathbf{K}\mathbf{H}^{\dagger}
\end{equation}

Where $\mathbf{H}(r)$ is the total magnetic field of the ellipsoids, $\mathbf{H}_0$ is the inducing magnetic field in each direction,
$\mathbf{K}$ is the susceptibility tensor and $\mathbf{H}^{\dagger}$ is the resultant uniform magnetic field at any point within
the ellipsoids. \cite{takahashi_ellipsoids_2017}  This approach is particularly useful for modeling spheroidal or ellipsoidal particles and is widely 
used in magnetorheological systems, chain-forming ferrofluids, and dense suspensions where particle-particle interactions are shape-dependent.

@@

The Volume of Fluid (VOF) method is a sharp-interface technique used to track the evolution of fluid-fluid interfaces within a fixed computational grid 
\cite{gopala_volume_2008, fleckenstein_volume-fluid-based_2015, deising_direct_2018}. In VOF, a scalar field representing the volume fraction of one 
fluid is defined for each grid cell to indicate the local phase composition. Interface reconstruction algorithms are then employed to geometrically 
localize the boundary between phases. The method is inherently mass-conserving and can handle large interface deformations, making it particularly 
effective for high-Reynolds-number flows and surface-tension-driven phenomena. However, VOF methods can struggle with accurately capturing interfacial 
curvature and are not well-suited for diffuse interfaces, which limits their applicability in modeling spinodal decomposition and phase separation 
processes such as those that form bijels.

Phase field methods use a diffuse-interface approach to model multiphase systems, representing fluid composition or phase identity through a continuous 
order parameter field governed by equations such as the Cahn-Hilliard or Allen-Cahn equations 
\cite{mendoza_evolution_2006, carmack_tuning_2018, chan_channel_2012}. These methods naturally capture smooth transitions between phases, enable
accurate computation of interfacial forces, and handle topological changes like coalescence and breakup with ease. Derived from variational principles, 
phase field models are thermodynamically consistent, making them particularly well-suited for simulating spinodal decomposition, wetting phenomena, and 
interfacial dynamics. However, their main limitations include the need for fine spatial resolution near interfaces to maintain accuracy and the relatively 
high computational cost of solving fourth-order PDEs, especially when coupled with full hydrodynamic models.

Brownian dynamics is a particle-based simulation method used to model the stochastic motion of colloidal particles suspended in a fluid 
\cite{huber_brownian_2019, yip_brownian_2005, elsawy_utility_2025}. It approximates the over-damped limit of Langevin dynamics by integrating 
particle trajectories under the influence of both deterministic forces—such as interparticle potentials or external fields—and random thermal 
forces representing collisions with solvent molecules. Hydrodynamic interactions may be included through approximate mobility tensors or omitted 
entirely, depending on the desired level of fidelity. Brownian dynamics is particularly well-suited for studying self-assembly, aggregation, and 
diffusion in dilute suspensions, where particle motion is dominated by thermal fluctuations. However, it does not model the surrounding fluid 
explicitly, limiting its ability to capture bulk fluid flow or interface-mediated interactions. As a result, Brownian dynamics is not sufficient 
on its own for simulating the coupled fluid-particle dynamics involved in bijel formation.

@@

In summary, the Volume of Fluid method offers robust interface tracking and mass conservation but lacks the ability to naturally capture interfacial curvature. 
Phase field methods provide thermodynamically consistent, diffuse-interface models ideal for phase separation, though they require fine resolution and are computationally intensive. 
Brownian dynamics excels at modeling particle-level stochastic motion but cannot resolve bulk fluid dynamics or interface evolution on its own. In contrast, the Lattice Boltzmann Method 
combines the strengths of mesoscopic fluid modeling with efficient handling of complex interfaces and straightforward coupling to particle dynamics. Its balance between computational 
efficiency, scalability, and physical fidelity across multiple scales makes LBM the most suitable method for simulating the rich interplay of hydrodynamics, phase separation, and particle 
interactions in stimuli-responsive emulsions such as bijels.

@@

The IT model employs level-set or front-tracking techniques to explicitly track the location of fluid interfaces by evolving a separate indicator 
function, commonly defined using a level set technique, alongside the LB solver. \cite{haghani_hassan_abadi_conservative_2018, liang_lattice_2023}
This method achieves high accuracy in resolving interface position and curvature, making it ideal for problems 
involving surface tension-driven flows or sharp interfacial features. However, it is computationally intensive as the indicator function must be implemented alongside
the LB solver.

@@

The density and velocity of the fluid at each grid point is calculated from the zeroth and first moment of the particle density
distribution,

\begin{equation}
    \begin{split}
        \rho &= \sum_i f_i \\
        \rho\vec{u} &= \sum_i f_i \mathbf{c_i}
    \end{split}
\end{equation}

The momentum flux tensor is calculated from the Chapman-Enskog expansion of \ref{eq:LBM_BGK} to the first order yielding,

\begin{equation}
    \begin{split}
        \rho\frac{\partial \vec{v}}{\partial t} + \nabla \cdot \Pi &= 0 \\
        \Pi_{\alpha \beta} &= \sum_i \mathbf{c_i}_\alpha \mathbf{c_i}_\beta (f_i^{eq} + (1 - \frac{1}{2 \tau})f_i^1)
    \end{split}
\end{equation}

The momentum flux tensor can be further expressed as a function of macroscopic variables,

\begin{equation}
    \begin{split}
        \Pi^{(0)}_{\alpha \beta} &= \sum_i \mathbf{c_i}_\alpha \mathbf{c_i}_\beta f_i^{eq} = p \delta_{\alpha \beta} + \rho \vec{u}_\alpha \vec{u}_\beta \\
        \Pi^{(1)}_{\alpha \beta} &= (1 - \frac{1}{2 \tau})\sum_i \mathbf{c_i}_\alpha \mathbf{c_i}_\beta f_i^{eq} = \nu (\nabla_{\alpha}(\rho \vec{\beta}) + \nabla_{\beta}(\rho \vec{\alpha}))
    \end{split}
\end{equation}

With $p = \rho c_s^2$ and $\delta_{\alpha \beta}$ the kronecker delta.

### 2025-04-08 chapters/final_remarks.tex
Changes 1. Added new lit review 2. Refined introduction and LBM methods 3. Fixed all references so that they have journals

Traditional bijel fabrication relies on thermally induced phase separation (TIPS), a method that, while effective, does not 
support continuous production. More recently developed techniques—such as Solvent Transfer Induced Phase Separation (STrIPS) 
and Vapor Induced Phase Separation (VIPS)—enable continuous fabrication and offer greater control over microstructure. However, 
these methods typically require modifications to the initial emulsion formulation, coupling the casting mixture to the resulting 
morphology. Decoupling these parameters would offer significantly greater flexibility in bijel design, allowing for independent 
tuning of material properties and processing conditions.

@@

This work utilized 2 ellipsoidal particle geometries chosen for comparisons to previous literature using this 
particle geometry. Particles based on cellulose nanocrystals or graphene nanoplates are now in use to fabricate
particle stabilized emulsions. These particles have also been shown to have capillary bridging and particle stacking,
not seen in the particles used in this work. These particles in bulk have been shown to have intrinsic ordering
that can be predicted using onsager theory. An investigation into how using rods or plate like particles would be 
instructive in identifying if onsager theory can be used in bijel formation and its link to bijel microstructure. 
\cite{tan_2d_2021}

Colloidal systems made with cohesive and soft particles have been shown in the literature to have different glass 
transition points and rheological behavior from their hard sphere counterparts. In bijels, attractive particles in 
particular have been suggested as a means to create "armored" bijels to improve their performance in catalytic materials, 
allowing higher flow rates to be used. Investigations into soft particles are of interest in biomedical
applications, with newly developed nanogels and hydrogels being suggested as drug carriers or vectors for stimuli 
response through temperature or pH. LBM methods that implement the immersed boundary method can be used to model soft
particles with a DLVO potential used to model electrostatics between particles. \cite{silva_lattice_2024}

Another avenue of exploration would be the use of gradient or rotating magnetic fields in place of the constant magnetic 
fields used here. One study on bijel microstructure showed that a gradient in the particle volume fraction can be used to 
create a gradient in the eventual domain size. A gradient field may be able to generate nematic order 
parameter gradients, affecting the particle packing of the bijel at different heights and varying the pore size as a function of 
height in the field gradient direction. In a ferrofluid or solution of magnetic colloids, a rotating field can assemble 
particles into chains or rings. In the context of bijel structural response, this can be used to tune bijel microstructure 
as this process can control the unjamming and rejamming of the particle monolayer, allowing for greater control over the 
resulting bijel microstructure than a constant field would have. In these simulations, the frequency of rotation would 
likely need to be tuned based upon how quickly particles respond to field in the bijel.

% Extensions to this work can be accomplished by investigating the effect of applying a magnetic field on the bijel 
% while under shear. Ferrofluid models that predict bingham plastic like flows, $\frac{\eta}{eta_{f}} = 1 + \frac{Mn^{*}(\phi_p)}{Mn}$, 
% have been developed and defined using the Mason number, $Mn = \frac{8\eta_{f} \dot{\gamma}}{\mu_{0} \mu_{f} \beta^{2} H_0^2}$. 
% \textcolor{blue}{https://doi.org/10.1122/1.4935850, https://linkinghub.elsevier.com/retrieve/pii/S1359029405000385}

>>>>

Traditional bijel fabrication relies on thermally induced phase separation (TIPS), a method that, while effective, does not 
support continuous production. More recently developed techniques such as Solvent Transfer Induced Phase Separation (STrIPS) 
and Vapor Induced Phase Separation (VIPS) enable continuous fabrication and offer greater control over microstructure. However, 
these methods typically require modifications to the initial emulsion formulation, coupling the casting mixture to the resulting 
morphology. Decoupling these parameters would offer significantly greater flexibility in bijel design, allowing for independent 
tuning of material properties and processing conditions.

@@

Building on the observed coupling between interfacial particle dynamics and bijel microstructure, we propose extending 
this work by investigating bijels stabilized with non-rigid, cohesive, or attractive particles. Unlike rigid ellipsoids, 
soft and cohesive particles are known to exhibit distinct glass transition behaviors and rheological responses compared 
to hard-sphere systems. \cite{weeks_introduction_2017, torquato_jammed_2010} In bijel architectures, attractive interactions have been 
proposed as a strategy to form “armored” interfaces, potentially enhancing mechanical integrity and enabling higher flow rates in catalytic applications. 
\cite{boakye-ansah_controlling_2020}
Similarly, soft particles such as nanogels and hydrogels have attracted interest in biomedical contexts, offering tunable responsiveness 
to environmental stimuli like temperature and pH—critical features for controlled drug delivery. Incorporating such particles 
could significantly alter jamming dynamics, interfacial rearrangement, and field-induced responsiveness. These effects may be 
explored computationally using Lattice Boltzmann Methods integrated with immersed boundary frameworks, incorporating attractive
particle  interactions to account for electrostatic cohesion and deformability \cite{silva_lattice_2024}. This direction opens opportunities 
to design bijel systems with enhanced adaptability and function, tailored for application-specific demands.

The current simulations, while able to capture structural evolution under external fields, do not fully resolve the complex dynamical landscape of the interfacial 
particle monolayer. In particular, they lack the resolution to distinguish between fast particle rearrangements and slow collective dynamics near the glass transition, 
which are known to govern the jamming and unjamming behavior of dense colloidal systems. \cite{weeks_introduction_2017, torquato_jammed_2010} 
As such, the interplay between interfacial ordering, particle-scale dynamics, and existing order can be characterized more convincingly by investigating the dynamics of
the particle more systematically by accounting for thermal fluctuations of the fluid. This would involve implementation of a fluctuating Lattice Boltzmann Method
which would facilitate identification of the glassy dynamics in the system, including the onset of arrest and structural aging characterized with the microstructure
in Gunther et al. \cite{gunther_timescales_2014}

Another promising avenue for future investigation involves the use of gradient or rotating magnetic fields in place of constant fields. This
would impart a force on the particles in addition to the torque that magnetic particles experience under constant fields. \cite{spatafora-salazar_hierarchical_2021, martinezpedrero_collective_2020}
Prior work has demonstrated that spatial gradients in particle volume fraction can generate domain size gradients in bijels, enabling spatial 
control over porosity and connectivity \cite{french_bicontinuous_2022}. Analogously, applying a magnetic field gradient may induce vertical 
gradients in nematic order, influencing local particle packing and yielding position-dependent pore structures across the bijel. This could be 
particularly useful for creating functionally graded materials with hierarchical transport properties.

In parallel, rotating magnetic fields offer a dynamic mechanism for restructuring particle arrangements at the interface. 
\cite{spatafora-salazar_hierarchical_2021, martinezpedrero_collective_2020}
In ferrofluids and magnetic colloidal suspensions, such fields have been shown to induce chain or ring formation, depending on the rotation frequency and field 
strength. \cite{spatafora-salazar_hierarchical_2021, martinezpedrero_collective_2020}
Translating this mechanism to bijels, a rotating field could be used to modulate interfacial jamming behavior, promoting controlled 
unjamming and reorientation of particles prior to re-jamming. This dynamic control could enable finer tuning of microstructure than static fields, 
potentially achieving reversible and spatially patterned domain morphologies. Importantly, the effectiveness of this strategy would depend on 
matching the rotation frequency to the particle relaxation timescales within the bijel, a factor that can be systematically explored in simulation.

### 2025-04-08 chapters/introduction_new.tex
Changes 1. Added new lit review 2. Refined introduction and LBM methods 3. Fixed all references so that they have journals

When two partially miscible fluids are mixed, they tend to phase separate into distinct domains to minimize the thermodynamic penalty associated with interfacial formation. 
To inhibit this process, stabilizers such as small-molecule surfactants, particles, or biopolymers are commonly introduced. Surfactants, for example, reduce the surface tension 
between the dispersed and continuous phases, thereby lowering the interfacial energy and promoting the stability of emulsions. A familiar example is soap, which facilitates the 
emulsification of dirt into droplets suspended in water through mechanical agitation.
Without such additives, the dispersed phase tends to coalesce in order to reduce interfacial area. The effectiveness of stabilizers can be characterized through the Pieranski 
model, $G_{ads} = \sigma A_{rm} (1 - \cos{\theta_c})^2$, where $G_{ads}$ is the free energy reduction upon particle adsorption, $\sigma$ is the surface tension between the two 
fluids, and $\theta_c$ is the contact angle of the stabilizing particle. While surfactants and biopolymers reduce interfacial tension, particle stabilizers function by adsorbing 
at the interface and effectively reducing the interfacial area between fluid domains.

Compared to surfactant-stabilized emulsions, particle-stabilized emulsions (Pickering emulsions) exhibit enhanced resistance to coalescence and coarsening, prompting renewed interest 
in their use across various applications. This resurgence is further driven by increasing environmental and toxicity concerns associated with small-molecule surfactants 
\cite{kaczerewska_environmental_2020, lechuga_acute_2016}. In contrast, particle stabilizers lower toxicity and improved biocompatibility, in addition to 
sustainable sourcing from particles based on cellulose or chitin. \cite{fujisawa_nanocellulose-stabilized_2017, tang_stimuli-responsive_2016, kalliola_carboxymethyl_2018}.
The microstructure of Pickering emulsions has been extensively studied, notably by Binks and Lumsdon in the early 2000s. Their investigations revealed that the emulsion droplet radius 
follows the relationship $R_e \propto \frac{\phi_w}{\phi_p}$, where $\phi_w$ and $\phi_p$ represent the volume fractions of water and particles, respectively \cite{binks_pickering_2001}. 
They also identified key parameters influencing microstructure, such as fluid composition and particle wettability. Neutrally wetting particles tend to stabilize nearly spherical droplets, 
while non-neutrally wetting particles can induce curvature, leading to the formation of bridged droplets, capillary aggregates, and other anisotropic structures, as illustrated in 
Figure~\ref{fig:state_diagram_particle_emulsions}.

@@

The dynamics of ellipsoid-stabilized bijels are distinct from their spherical counterparts, displaying two characteristic timescales. Initially, particles adsorb and align with the interface, 
followed by a slower reorganization phase driven by capillary interactions between particles \cite{gunther_timescales_2014}. Experimental results confirm that the domain size in rod- and ellipsoid-stabilized 
bijels still follows the scaling law $L \propto \frac{1}{\phi_p}$, though the proportionality constant depends on particle geometry \cite{hijnen_bijels_2015, madivala_exploiting_2009, daware_emulsions_2015}.

>>>>

When two partially miscible fluids are mixed, they tend to phase separate into distinct domains to minimize the thermodynamic penalty associated with interfacial formation. 
To inhibit this process, stabilizers such as small-molecule surfactants, particles, or biopolymers are commonly introduced. Surfactants, for example, reduce the surface tension 
between the dispersed and continuous phases, thereby lowering the interfacial energy and promoting the stability of emulsions. A familiar example is soap, which facilitates the 
emulsification of dirt into droplets suspended in water through mechanical agitation.
Without such additives, the dispersed phase tends to coalesce in order to reduce interfacial area. The effectiveness of stabilizers can be characterized through the Pieranski 
model, $G_{ads} = \sigma A_{rm} (1 - \cos{\theta_c})^2$, where $G_{ads}$ is the free energy reduction upon particle adsorption, $\sigma$ is the surface tension between the two 
fluids, and $\theta_c$ is the contact angle of the stabilizing particle. While surfactants and biopolymers reduce interfacial tension, particle stabilizers function by adsorbing 
at the interface and effectively reducing the interfacial area between fluid domains. The free energy associated with the decrease in interfacial area $G_{ads}$ can be orders of magnitude
larger than the thermal energy of the system $k_b T$ which makes the particle irreversibly adsorbed on the interface. \cite{ngai_particle-stabilized_2015}

Compared to surfactant-stabilized emulsions, particle-stabilized emulsions (Pickering emulsions) exhibit enhanced resistance to coalescence and coarsening, prompting renewed interest 
in their use across various applications. \cite{ngai_particle-stabilized_2015} 
This resurgence is further driven by increasing environmental and toxicity concerns associated with small-molecule surfactants 
\cite{kaczerewska_environmental_2020, lechuga_acute_2016}. In contrast, particle stabilizers lower toxicity and improved biocompatibility, in addition to 
sustainable sourcing from particles based on cellulose or chitin. \cite{fujisawa_nanocellulose-stabilized_2017, tang_stimuli-responsive_2016, kalliola_carboxymethyl_2018}.
The microstructure of Pickering emulsions has been extensively studied, notably by Binks and Lumsdon in the early 2000s. Their investigations revealed that the emulsion droplet radius 
follows the relationship $R_e \propto \frac{\phi_w}{\phi_p}$, where $\phi_w$ and $\phi_p$ represent the volume fractions of water and particles, respectively \cite{binks_pickering_2001}. 
They also identified key parameters influencing microstructure, such as fluid composition and particle wettability. Neutrally wetting particles tend to stabilize nearly spherical droplets, 
while non-neutrally wetting particles can induce curvature, leading to the formation of bridged droplets, capillary aggregates, and other anisotropic structures, as illustrated in 
Figure~\ref{fig:state_diagram_particle_emulsions}.

@@

The dynamics of ellipsoid-stabilized bijels are distinct from their spherical counterparts, displaying two characteristic timescales. Initially, particles adsorb and align with the interface, 
followed by a slower reorganization phase driven by capillary interactions between particles \cite{gunther_timescales_2014}. Experimental results confirm that the domain size in rod and ellipsoid stabilized 
bijels still follows the scaling law $L \propto \frac{1}{\phi_p}$, though the proportionality constant depends on particle geometry \cite{hijnen_bijels_2015, madivala_exploiting_2009, daware_emulsions_2015}.

### 2025-04-08 chapters/paper1_new.tex
Changes 1. Added new lit review 2. Refined introduction and LBM methods 3. Fixed all references so that they have journals

@@

Figure~\ref{fig:rdf} shows the radial distribution function (RDF) at time steps \(t = 1000\Delta t,\ 10000\Delta t,\ 100000\Delta t\). 
For ellipsoidal particles, characteristic contact distances include \(2R_{\parallel}\), \(2R_{\perp}\), and \(R_{\parallel} + R_{\perp}\), 
indicated by dashed lines in the figure. The largest RDF peak for both oblate and spherical particles occurs at \(2R_\perp\), suggesting 
contact around the particle circumference, while prolate particles preferentially align side-by-side. These configurations enable denser 
packing at the interface.

In the early stages of the simulation, RDF peaks are less distinct due to the more random particle orientations. Notably, for prolate particles, 
the RDF shows an initial dip at \(2R_\parallel\), indicating that tip-to-tip arrangements are rare. As time progresses, the peak heights increase, 
particularly under magnetic fields, supporting the idea that magnetic alignment promotes side-by-side and tip-to-tip ordering. For oblate particles, 
a shoulder emerges near \(R_\parallel + R_\perp\), indicating occasional tilting out of the interface to reduce interfacial coverage. In contrast, 
for prolate particles, the dip at \(2R_\parallel\) disappears over time, suggesting increased tip-to-tip alignment—consistent with field-driven 
particle orientation that supports close packing along both particle axes. This layered structure is also visible in the snapshots in 
Figure~\ref{fig:packing_viz}.

>>>>

This section is based on work published in the Royal Society of Chemistry's Soft Matter journal. \cite{karthikeyan_formation_2024} 
It is reproduced from Ref \cite{karthikeyan_formation_2024} with permission from the Royal Society of Chemistry.

@@

Figure~\ref{fig:rdf} shows the radial distribution function (RDF) at time steps $t = 1000\Delta t, 10000\Delta t, 100000\Delta t$. 
For ellipsoidal particles, characteristic contact distances include \(2R_{\parallel}\), \(2R_{\perp}\), and \(R_{\parallel} + R_{\perp}\), 
indicated by dashed lines in the figure. The largest RDF peak for both oblate and spherical particles occurs at \(2R_\perp\), suggesting 
contact around the particle circumference, while prolate particles preferentially align side-by-side. These configurations enable denser 
packing at the interface.

In the early stages of the simulation, RDF peaks are less distinct due to the more random particle orientations. Notably, for prolate particles, 
the RDF shows an initial dip at \(2R_\parallel\), indicating that tip-to-tip arrangements are rare. As time progresses, the peak heights increase, 
particularly under magnetic fields, supporting the idea that magnetic alignment promotes side-by-side and tip-to-tip ordering. For oblate particles, 
a shoulder emerges near \(R_\parallel + R_\perp\), indicating occasional tilting out of the interface to reduce interfacial coverage. In contrast, 
for prolate particles, the dip at \(2R_\parallel\) disappears over time, suggesting increased tip-to-tip alignment, consistent with field-driven 
particle orientation that supports close packing along both particle axes. This layered structure is also visible in the snapshots in 
Figure~\ref{fig:packing_viz}.
